---
title: "ANA515_Final Project"
author: "Thi Huyen Nguyen"
date: "12/12/2021"
output: html_document
---
# Goal of the project: Use visualizations to analyze how time affected customer trips and how various users made trips from different bases.

# Installing Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


install.packages("tidyverse")
install.packages("ggplot2")
install.packages("ggthemes")
install.packages("lubridate")
install.packages("dplyr")
install.packages("tidyr")
install.packages("DT")
install.packages("scales")


# calling installed packages
```{r, include=FALSE}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
```

# Creating vector of colors to be implemented in our plots
```{r, include=FALSE}
colors = c("#CC1011", "#665555", "#05a399", "#cfcaca", "#f5e840", "#0683c9", "#e075b0")
```

#Using read.csv to import data from "C:/Users/16087/OneDrive/Documents", 
```{r, include=FALSE}
apr_data <- read.csv("uber-raw-data-apr14.csv")
may_data <- read.csv("uber-raw-data-may14.csv")
jun_data <- read.csv("uber-raw-data-jun14.csv")
jul_data <- read.csv("uber-raw-data-jul14.csv")
aug_data <- read.csv("uber-raw-data-aug14.csv")
sep_data <- read.csv("uber-raw-data-sep14.csv")
```

# Using summary to describe datasets
```{r, include=FALSE}
summary(apr_data)
summary(may_data)
summary(jun_data)
summary(jul_data)
summary(aug_data)
summary(sep_data)
```

# Using rbind to bind dataframes
```{r, include=FALSE}
data_2014 <- rbind(apr_data,may_data, jun_data, jul_data, aug_data, sep_data)
```

# Preparing data by Converting and Creating factors of time objects
```{r, include=FALSE}
data_2014$Date.Time <- as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S")
data_2014$Time <- format(as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
data_2014$Date.Time <- ymd_hms(data_2014$Date.Time)
data_2014$day <- factor(day(data_2014$Date.Time))
data_2014$month <- factor(month(data_2014$Date.Time, label = TRUE))
data_2014$year <- factor(year(data_2014$Date.Time))
data_2014$dayofweek <- factor(wday(data_2014$Date.Time, label = TRUE))
data_2014$hour <- factor(hour(hms(data_2014$Time)))
data_2014$minute <- factor(minute(hms(data_2014$Time)))
data_2014$second <- factor(second(hms(data_2014$Time)))
```

#Plotting the trips by the hours in a day: Tnumber of trips are higher in the evening around 5:00 and 6:00 PM.
```{r, include=FALSE}
hour_data <- data_2014 %>%
           group_by(hour) %>%
               dplyr::summarize(Total = n()) 
datatable(hour_data)
ggplot(hour_data, aes(hour, Total)) + 
        geom_bar( stat = "identity", fill = "steelblue", color = "red") +
           ggtitle("Trips Every Hour") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
month_hour <- data_2014 %>%
          group_by(month, hour) %>%
             dplyr::summarize(Total = n())
ggplot(month_hour, aes(hour, Total, fill = month)) + 
       geom_bar( stat = "identity") +
          ggtitle("Trips by Hour and Month") +
           scale_y_continuous(labels = comma)
```

# Plotting data by trips during every day of the month: number of trips are higher in the day 30th of the month.
```{r, include=FALSE}
day_group <- data_2014 %>%
          group_by(day) %>%
             dplyr::summarize(Total = n()) 
datatable(day_group)
ggplot(day_group, aes(day, Total)) + 
        geom_bar( stat = "identity", fill = "steelblue") +
           ggtitle("Trips Every Day") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma)
```

# Plotting data by trips by day and month: number of trips are higher in day 30 and September
```{r, include=FALSE}
day_month_group <- data_2014 %>%
         group_by(month, day) %>%
             dplyr::summarize(Total = n())

ggplot(day_month_group, aes(day, Total, fill = month)) + 
        geom_bar( stat = "identity") +
           ggtitle("Trips by Day and Month") +
            scale_y_continuous(labels = comma) +
            scale_fill_manual(values = colors)
```


# Ploting Number of Trips taking place during months in a year: September is the most busy month of the year in the data
```{r, include=FALSE}
month_group <- data_2014 %>%
          group_by(month) %>%
             dplyr::summarize(Total = n()) 
datatable(month_group)

ggplot(month_group, aes(month, Total, fill = month)) + 
        geom_bar( stat = "identity") +
           ggtitle("Trips by Month") +
            theme(legend.position = "none") +
            scale_y_continuous(labels = comma) +
            scale_fill_manual(values = colors)
```

# Ploting trip volume by Base: number of trips are higher in Base 02617
```{r, include=FALSE}
ggplot(data_2014, aes(Base)) + 
 geom_bar(fill = "darkred") +
 scale_y_continuous(labels = comma) +
 ggtitle("Trips by Bases")
```
# Ploting trips by Bases and Month: number of trips are higher Base 02617 and September
```{r, include=FALSE}
ggplot(data_2014, aes(Base, fill = month)) + 
 geom_bar(position = "dodge") +
 scale_y_continuous(labels = comma) +
 ggtitle("Trips by Bases and Month") +
 scale_fill_manual(values = colors)
```

# Ploting trip volume by Bases and Day of Week: number of trips are higher in Base 02617 and on Thursday.
```{r, include=FALSE}
ggplot(data_2014, aes(Base, fill = dayofweek)) + 
 geom_bar(position = "dodge") +
 scale_y_continuous(labels = comma) +
 ggtitle("Trips by Bases and DayofWeek") +
 scale_fill_manual(values = colors)
```
# Creating a Heatmap visualization of day, hour: number of trips are higher on the day 16th of the month and from 5PM to 6PM.
```{r, include=FALSE}
day_and_hour <- data_2014 %>%
         group_by(day, hour) %>%
            dplyr::summarize(Total = n())
ggplot(day_and_hour, aes(day, hour, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Hour and Day")
```
# Creating a Heatmap by month and bases
```{r, include=FALSE}
month_base <-  data_2014 %>%
                    group_by(Base, month) %>%
                        dplyr::summarize(Total = n()) 
day0fweek_bases <-  data_2014 %>%
                    group_by(Base, dayofweek) %>%
                        dplyr::summarize(Total = n()) 
ggplot(month_base, aes(Base, month, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Month and Bases")
```

# Creating heatmap by Bases and Day of the week: number of trips are higher in Base 02617 and on Thursday.
```{r, include=FALSE}
ggplot(day0fweek_bases, aes(Base, dayofweek, fill = Total)) +
            geom_tile(color = "white") +
              ggtitle("Heat Map by Bases and Day of Week")
```

# Creating a map visualization of rides in New York
```{r, include=FALSE}
min_lat <- 40.5774
max_lat <- 40.9176
min_long <- -74.15
max_long <- -73.7004
ggplot(data_2014, aes(x=Lon, y=Lat)) +
  geom_point(size=1, color = "blue") +
     scale_x_continuous(limits=c(min_long, max_long)) +
      scale_y_continuous(limits=c(min_lat, max_lat)) +
        theme_map() +
           ggtitle("NYC MAP BASED ON UBER RIDES DURING 2014 (APR-SEP)")
ggplot(data_2014, aes(x=Lon, y=Lat, color = Base)) +
  geom_point(size=1) +
     scale_x_continuous(limits=c(min_long, max_long)) +
      scale_y_continuous(limits=c(min_lat, max_lat)) +
       theme_map() +
          ggtitle("NYC MAP BASED ON UBER RIDES DURING 2014 (APR-SEP) by BASE")
```
